Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-05-24 ms.tex
Add full posterior
Clean up some marginpar formatting

If we were to suppress the time parameter $\tau$, then we could reduce the likelihood to the form
%
\begin{equation}
    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{i \gamma} \rho \right|^2 \right]
\end{equation}
%
for a complex scalar measurement $z$. This likelihood also describes the distribution of the length $\rho$ and angle $\gamma$ of a vector whose Cartesian components are perturbed with normally distributed errors. For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.
%
\marginpar{Picture here?}

\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}
%
We know that the matched\nobreakdashes-filter estimator's distribution, given a particular true value of $\rho$, $\gamma$, and $\tau$, must go to zero at $\hat{\rho}=0$ and must be periodic in $\gamma$ but not $\tau$. From just these properties, the estimator's distribution should be an approximate Gaussian, but wrapped diagonally around a cylinder where the radius represents $\hat{\rho}$, the angle around the cylinder represents $\hat{\gamma}$, and the distance up or down the cylinder represents $\hat{\tau}$.

There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.
%
\marginpar{Give some examples.}
%
An alternative is to just discard the one periodic parameter, $\gamma$, entirely. We do not know the exact distribution of the remaining parameters $\rho$ and $\tau$, but we are less liable to make egregious errors by assuming that they are Gaussian than if we assumed that the estimation errors in $\gamma$ were Gaussian. Projecting out $\gamma$ gives us the covariance matrix
%
\begin{equation}
    \cov \left(
        \begin{array}{c}
            \hat{\rho} - \rho \\
            \hat{\tau} - \tau
        \end{array}
    \right) =
    \left(
        \begin{array}{cc}
            1 & 0 \\
            0 & \rho^2/{\omega_\mathrm{rms}}^2
        \end{array}
    \right).
\end{equation}

@@

\marginpar{Explain why none of these expressions depend on the coalescence phase $\phi_c$.}%
%
If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, x, \psi) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, x, \psi
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $u$, and polarization angle $\psi$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
    \right]
\end{equation}
%
\marginpar{Cite \ac{GRB} prospects paper appendix; a very similar equation appears there. Is there somewhere else that has this formula? Shutz doesn't, and Finn \& Chernoff's notation is too heavy.}%
%
It is easily shown from \Eref{eq:full-signal-model} that
%
\begin{eqnarray}\label{eq:snr-model}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left\{
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 u^2 + u^4 \right) \nonumber\\
            &+ \left(1 - u^2 \right)^2 \left[\left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {4\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {4\psi} \right]
        \right\}.
\end{eqnarray}

@@

\subsection{Marginalization}

@@

>>>>

If we were to suppress the time parameter $\tau$, then we could reduce the likelihood to the form
%
\begin{equation}
    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{i \gamma} \rho \right|^2 \right]
\end{equation}
%
for a complex scalar measurement $z$. This likelihood also describes the distribution of the length $\rho$ and angle $\gamma$ of a vector whose Cartesian components are perturbed with normally distributed errors. For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.%
%
\marginpar{Picture here?}

\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}%
%
We know that the matched\nobreakdashes-filter estimator's distribution, given a particular true value of $\rho$, $\gamma$, and $\tau$, must go to zero at $\hat{\rho}=0$ and must be periodic in $\gamma$ but not $\tau$. From just these properties, the estimator's distribution should be an approximate Gaussian, but wrapped diagonally around a cylinder where the radius represents $\hat{\rho}$, the angle around the cylinder represents $\hat{\gamma}$, and the distance up or down the cylinder represents $\hat{\tau}$.

There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.%
%
\marginpar{Give some examples.}%
%
An alternative is to just discard the one periodic parameter, $\gamma$, entirely. We do not know the exact distribution of the remaining parameters $\rho$ and $\tau$, but we are less liable to make egregious errors by assuming that they are Gaussian than if we assumed that the estimation errors in $\gamma$ were Gaussian. Projecting out $\gamma$ gives us the covariance matrix
%
\begin{equation}
    \cov \left(
        \begin{array}{c}
            \hat{\rho} - \rho \\
            \hat{\tau} - \tau
        \end{array}
    \right) =
    \left(
        \begin{array}{cc}
            1 & 0 \\
            0 & \rho^2/{\omega_\mathrm{rms}}^2
        \end{array}
    \right).
\end{equation}

@@

\marginpar{Explain why none of these expressions depend on the coalescence phase $\phi_c$.}%
%
If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, u, \psi) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, u, \psi
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $u$, and polarization angle $\psi$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
    \right]
\end{equation}
%
\marginpar{Cite \ac{GRB} prospects paper appendix; a very similar equation appears there. Is there somewhere else that has this formula? Shutz doesn't, and Finn \& Chernoff's notation is too heavy.}%
%
It is easily shown from \Eref{eq:full-signal-model} that
%
\begin{eqnarray}\label{eq:snr-model}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left\{
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 u^2 + u^4 \right) \nonumber\\
            &+ \left(1 - u^2 \right)^2 \left[\left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {4\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {4\psi} \right]
        \right\}.
\end{eqnarray}

@@

\subsection{Marginalization over $t_\oplus$}

@@

\subsection{Posterior}

\begin{eqnarray}
    p(
		\mathbf{n} |
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N) \nonumber\\
    &= p(\mathbf{n} |
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N) \nonumber\\
        &\times
        p(\mathbf{n} |
		\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N
    )
\end{eqnarray}
%
where
%
\begin{equation*}
	\fl
    p(\mathbf{n} |
	\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N)
	\propto
	\int_0^\frac{\pi}{2}
	\int_0^1
	\int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}}
	\exp \left[
	        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
	    \right]
	{D_\mathrm{L}}^m
	\,\mathrm{d}D_\mathrm{L}
	\,\mathrm{d}u
	\,\mathrm{d}\psi.
\end{equation*}
